Parallel Monte Carlo Synthetic Acceleration Methods for Discrete Transport Problems

### 2013-09-03 doc/dissertation/presentation/defense.tex
updating solution evolution slide

  \begin{itemize}
  \item Monte Carlo Synthetic Acceleration Methods
    \bigskip
  \item Application to Neutron Transport
    \bigskip
  \item Application to Fluid Flow
    \bigskip
  \item Parallelization of MCSA
    \bigskip
  \item Summary
  \end{itemize}

@@

  \begin{itemize}
  \item New algorithms required to leverage new computational
    resources
    \begin{itemize}
    \item Tremendous computational resources are required with
      $O(\sn{1}{9})$ element meshes and $O(100,000)+$ cores used today
      for neutronics and fluid problems (Evans,2010)(Pawlowski,2012)
    \end{itemize}
    \bigskip
  \item Necessary to determine applicability to reactor physics and
    potential performance improvements before moving forward with
    other work
    \bigskip
  \item Physics-driven development
    \begin{itemize}
    \item Research applicability and potential improvements to
      neutronics and fluid flow
    \item Offer solutions or a potential path forward for the observed
      issues
    \item Work to improve iterative and parallel performance where
      possible
    \end{itemize}
  \end{itemize}

@@

%%---------------------------------------------------------------------------%%
\begin{frame}{Research Outline}
  \begin{itemize}
  \item Development of a linear scheme for discrete systems leveraging
    Monte Carlo Synthetic Acceleration
    \medskip
    \begin{itemize}
    \item Application to neutron transport
    \item Research is required to study MCSA preconditioning
    \item Iterative performance is of concern
    \end{itemize}
    \bigskip
  \item Development of a nonlinear scheme for discrete systems
    leveraging Monte Carlo Synthetic Acceleration
    \medskip
    \begin{itemize}
    \item Monte Carlo is a more natural fit
    \item Application to fluid flow
    \item Convergence of the linear model is of concern
    \end{itemize}
    \bigskip
  \item Parallelization of Monte Carlo Synthetic Acceleration
    \medskip
    \begin{itemize}
    \item Parallel strategies taken from modern reactor physics
      methods
    \item Research is required to explore varying parallel strategies
    \item Parallel scalability is of concern
    \end{itemize}
  \end{itemize}
\end{frame}

@@

  \begin{itemize}
  \item \textbf{Monte Carlo Synthetic Acceleration Methods}
    \bigskip
  \item Application to Neutron Transport
    \bigskip
  \item Application to Fluid Flow
    \bigskip
  \item Parallelization of MCSA
    \bigskip
  \item Summary
  \end{itemize}

@@

  \begin{itemize}
  \item First proposed by J. Von Neumann and S.M. Ulam in the 1940's
    \medskip \medskip
  \item Earliest published reference in 1950
    \medskip \medskip
  \item General lack of published work
    \medskip \medskip
  \item Modern work by Evans and others has yielded new
    applications\let\thefootnote\relax\footnote{\tiny{Thomas Evans and
        Scott Mosher, "A Monte Carlo Synthetic Acceleration method for
        the non-linear, time-dependent diffusion equation", American
        Nuclear Society - International Conference on Mathematics,
        Computational Methods and Reactor Physics, 2009.}}
  \end{itemize}
\end{frame}

@@

  \begin{figure}[h!]
    \begin{center}
      \includegraphics[width=4in]{adjoint_1.png}
    \end{center}
    \caption{\textbf{Adjoint solution to Poisson Equation.}
      \textit{\sn{1}{0} total histories, 0.286 seconds CPU time.} }
  \end{figure}

  \begin{figure}[h!]
    \begin{center}
      \includegraphics[width=4in]{adjoint_10.png}
    \end{center}
    \caption{\textbf{Adjoint solution to Poisson Equation.}
      \textit{\sn{1}{1} total histories, 0.278 seconds CPU time.} }
  \end{figure}

  \begin{figure}[h!]
    \begin{center}
      \includegraphics[width=4in]{adjoint_100.png}
    \end{center}
    \caption{\textbf{Adjoint solution to Poisson Equation.}
      \textit{\sn{1}{2} total histories, 0.275 seconds CPU time.} }
  \end{figure}

  \begin{figure}[h!]
    \begin{center}
      \includegraphics[width=4in]{adjoint_1000.png}
    \end{center}
    \caption{\textbf{Adjoint solution to Poisson Equation.}
      \textit{\sn{1}{3} total histories, 0.291 seconds CPU time.} }
  \end{figure}

  \begin{figure}[h!]
    \begin{center}
      \includegraphics[width=4in]{adjoint_10000.png}
    \end{center}
    \caption{\textbf{Adjoint solution to Poisson Equation.}
      \textit{\sn{1}{4} total histories, 0.428 seconds CPU time.} }
  \end{figure}

  \begin{figure}[h!]
    \begin{center}
      \includegraphics[width=4in]{adjoint_100000.png}
    \end{center}
    \caption{\textbf{Adjoint solution to Poisson Equation.}
      \textit{\sn{1}{5} total histories, 1.76 seconds CPU time.} }
  \end{figure}

  \begin{figure}[h!]
    \begin{center}
      \includegraphics[width=4in]{adjoint_1000000.png}
    \end{center}
    \caption{\textbf{Adjoint solution to Poisson Equation.}
      \textit{\sn{1}{6} total histories, 15.1 seconds CPU time.} }
  \end{figure}

  \begin{figure}[h!]
    \begin{center}
      \includegraphics[width=4in]{adjoint_10000000.png}
    \end{center}
    \caption{\textbf{Adjoint solution to Poisson Equation.}
      \textit{\sn{1}{7} total histories, 149 seconds CPU time.} }
  \end{figure}

@@

  \begin{itemize}
  \item Monte Carlo Synthetic Acceleration Methods
    \bigskip
  \item \textbf{Application to Neutron Transport}
    \bigskip
  \item Application to Fluid Flow
    \bigskip
  \item Parallelization of MCSA
    \bigskip
  \item Summary
  \end{itemize}

@@

  \begin{itemize}
  \item Monte Carlo Synthetic Acceleration Methods
    \bigskip
  \item Application to Neutron Transport
    \bigskip
  \item \textbf{Application to Fluid Flow}
    \bigskip
  \item Parallelization of MCSA
    \bigskip
  \item Summary
  \end{itemize}

@@

  \begin{itemize}
  \item Monte Carlo Synthetic Acceleration Methods
    \bigskip
  \item Application to Neutron Transport
    \bigskip
  \item Application to Fluid Flow
    \bigskip
  \item \textbf{Parallelization of MCSA}
    \bigskip
  \item Summary
  \end{itemize}

@@

  \begin{itemize}
  \item Monte Carlo Synthetic Acceleration Methods
    \bigskip
  \item Application to Neutron Transport
    \bigskip
  \item Application to Fluid Flow
    \bigskip
  \item Parallelization of MCSA
    \bigskip
  \item \textbf{Summary}
  \end{itemize}

>>>>

  \begin{itemize}
  \item \textbf{Introduction}
    \bigskip
  \item Monte Carlo Synthetic Acceleration Methods
    \bigskip
  \item Application to Neutron Transport
    \bigskip
  \item Application to Fluid Flow
    \bigskip
  \item Parallelization of MCSA
    \bigskip
  \item Summary
  \end{itemize}

@@

  \begin{itemize}
  \item New algorithms required to leverage new computational
    resources
    \begin{itemize}
    \item Tremendous computational resources are required with
      $O(\sn{1}{9})$ element meshes and $O(100,000)+$ cores used today
      for neutronics and fluid problems (Evans,2010)(Pawlowski,2012)
    \end{itemize}
    \bigskip
  \item Necessary to determine applicability to reactor physics and
    potential performance improvements before moving forward with
    other work
    \bigskip
  \item Physics-driven development
    \begin{itemize}
    \item Research applicability and potential improvements to
      neutronics and fluid flow
    \item Offer solutions or a potential path forward for the observed
      issues
    \item Work to improve iterative and parallel performance
    \end{itemize}
  \end{itemize}

@@

%%---------------------------------------------------------------------------%%
\begin{frame}{Research Outline}
  \begin{itemize}
  \item Development of a linear scheme for the $SP_N$ equations
    leveraging Monte Carlo Synthetic Acceleration
    \medskip
    \begin{itemize}
    \item Application to neutron transport
    \item Research is required to study MCSA preconditioning
    \item Iterative performance is of concern
    \end{itemize}
    \bigskip
  \item Development of a nonlinear scheme for the Navier-Stokes
    equations leveraging Monte Carlo Synthetic Acceleration
    \medskip
    \begin{itemize}
    \item Monte Carlo is a more natural fit
    \item Application to fluid flow
    \item Convergence of the linear model is of concern
    \end{itemize}
    \bigskip
  \item Parallelization of Monte Carlo Synthetic Acceleration
    \medskip
    \begin{itemize}
    \item Parallel strategies taken from modern reactor physics
      methods
    \item Research is required to explore varying parallel strategies
    \item Parallel scalability is of concern
    \end{itemize}
  \end{itemize}
\end{frame}

@@

  \begin{itemize}
  \item Introduction
    \bigskip
  \item \textbf{Monte Carlo Synthetic Acceleration Methods}
    \bigskip
  \item Application to Neutron Transport
    \bigskip
  \item Application to Fluid Flow
    \bigskip
  \item Parallelization of MCSA
    \bigskip
  \item Summary
  \end{itemize}

@@

  \begin{itemize}
  \item First proposed by J. Von Neumann and S.M. Ulam in the 1940's
    \medskip \medskip
  \item Earliest published reference in 1950
    \medskip \medskip
  \item General lack of published work
    \medskip \medskip
  \item Modern work by Evans and others has yielded new
    applications\let\thefootnote\relax\footnote{\tiny{Thomas Evans and
        Scott Mosher, "A Monte Carlo Synthetic Acceleration method for
        the non-linear, time-dependent diffusion equation", American
        Nuclear Society - International Conference on Mathematics,
        Computational Methods and Reactor Physics, 2009.}}
  \end{itemize}

@@

  \begin{figure}[h!]
    \begin{center}
      \includegraphics<1>[width=4in]{adjoint_1.png}
      \includegraphics<2>[width=4in]{adjoint_10.png}
      \includegraphics<3>[width=4in]{adjoint_100.png}
      \includegraphics<4>[width=4in]{adjoint_1000.png}
      \includegraphics<5>[width=4in]{adjoint_10000.png}
      \includegraphics<6>[width=4in]{adjoint_100000.png}
      \includegraphics<7>[width=4in]{adjoint_1000000.png}
      \includegraphics<8>[width=4in]{adjoint_10000000.png}
    \end{center}
    \caption{
      \only<1>{\textbf{Adjoint solution to Poisson Equation.}
        \textit{\sn{1}{0} total histories, 0.286 seconds CPU time.} }
      \only<2>{\textbf{Adjoint solution to Poisson Equation.}
        \textit{\sn{1}{1} total histories, 0.278 seconds CPU time.} }
      \only<3>{\textbf{Adjoint solution to Poisson Equation.}
        \textit{\sn{1}{2} total histories, 0.275 seconds CPU time.} }
      \only<4>{\textbf{Adjoint solution to Poisson Equation.}
        \textit{\sn{1}{3} total histories, 0.291 seconds CPU time.} }
      \only<5>{\textbf{Adjoint solution to Poisson Equation.}
        \textit{\sn{1}{4} total histories, 0.428 seconds CPU time.} }
      \only<6>{\textbf{Adjoint solution to Poisson Equation.}
        \textit{\sn{1}{5} total histories, 1.76 seconds CPU time.} }
      \only<7>{\textbf{Adjoint solution to Poisson Equation.}
        \textit{\sn{1}{6} total histories, 15.1 seconds CPU time.} }
      \only<8>{\textbf{Adjoint solution to Poisson Equation.}
        \textit{\sn{1}{7} total histories, 149 seconds CPU time.} } 
    }
  \end{figure}

@@

  \begin{itemize}
  \item Introduction
    \bigskip
  \item Monte Carlo Synthetic Acceleration Methods
    \bigskip
  \item \textbf{Application to Neutron Transport}
    \bigskip
  \item Application to Fluid Flow
    \bigskip
  \item Parallelization of MCSA
    \bigskip
  \item Summary
  \end{itemize}

@@

  \begin{itemize}
  \item Introduction
    \bigskip
  \item Monte Carlo Synthetic Acceleration Methods
    \bigskip
  \item Application to Neutron Transport
    \bigskip
  \item \textbf{Application to Fluid Flow}
    \bigskip
  \item Parallelization of MCSA
    \bigskip
  \item Summary
  \end{itemize}

@@

  \begin{itemize}
  \item Introduction
    \bigskip
  \item Monte Carlo Synthetic Acceleration Methods
    \bigskip
  \item Application to Neutron Transport
    \bigskip
  \item Application to Fluid Flow
    \bigskip
  \item \textbf{Parallelization of MCSA}
    \bigskip
  \item Summary
  \end{itemize}

@@

  \begin{itemize}
  \item Introduction
    \bigskip
  \item Monte Carlo Synthetic Acceleration Methods
    \bigskip
  \item Application to Neutron Transport
    \bigskip
  \item Application to Fluid Flow
    \bigskip
  \item Parallelization of MCSA
    \bigskip
  \item \textbf{Summary}
  \end{itemize}
